---
title: "Ceilometer project - Analysis 3 - ARIMA models"
output: html_notebook
---


Section 0 - Various libs

```{r}
library(ggplot2)
library(forecast)
library(Metrics) #For various test metrics
library(dplyr)
library(dtw)
library(hydroGOF) #for IOA
library(tidyr)
library(tidyverse)
```

Section 1 - Data prep

Read data

```{r}
df_merriwa <- read.csv("D:/MSDS/DS Project/Git repos/group-18/Cleaned_Data/Merriwa_cleaned.csv")
df_lidcombe <- read.csv("D:/MSDS/DS Project/Git repos/group-18/Cleaned_Data/Lidcombe_cleaned.csv")

df_merriwa
```


Defining a universal train test split
```{r}
df_merriwa_test <- tail(df_merriwa, n=24)
df_lidcombe_test <- tail(df_lidcombe, n = 24)

df_merriwa_train <- head(df_merriwa, n = nrow(df_merriwa) - 24)
df_lidcombe_train <- head(df_lidcombe, n = nrow(df_lidcombe) - 24)
```

Section 1.2 Various test stats

```{r}
ioa <- function(sim, obs){
  mean_sim <- mean(sim)
  mean_obs <- mean(obs)
  
  ia <- 1 - (sum(sim - obs)^2)/(sum(abs(sim-mean_obs)+abs(obs-mean_obs))^2)
  
  return(ia)
}


mbe <- function(sim, obs){
  return(mean(sim - obs))
}


mase <- function(sim, obs, training_series){
  n <- length(training_series)
  d <- sum(abs(diff(training_series, lag = 1)))/(n-1)
  
  errors <- abs(sim - obs)
  
  return(mean(errors)/d)
  
}
```


```{r}
stats_metrics <- function(sim, obs, training_data, title){

  ioa <- ioa(sim, obs)
  rmse <- rmse(sim, obs)
  mae <- mae(sim, obs)
  mbe <- mbe(sim, obs)
  mase <- mase(sim, obs, training_data)
  dtw_dist <- dtw(sim, obs)$distance
  correlation <- cor(sim, obs)

  print(title)
  print('IOA RMSE MAE MBE DTW MASE CORR')
  sprintf("%0.5f %0.2f %0.2f %0.2f %0.2f %0.5f %0.5f", ioa, rmse, mae, mbe, dtw_dist, mase, correlation)
  }
```



Section 2 - Modelling

2.1 Simple auto ARIMA model - no external vars

Looking at just auto arima with no external vars - just at Merriwa for now

```{r}
arima_merriwa <- auto.arima(df_merriwa_train[,'CL_BLH'])

arima_merriwa
checkresiduals(arima_merriwa)
```
Simple forecast - no external factors

```{r}
df_merriwa_forecast <- forecast(arima_merriwa, h = 24)


plot(df_merriwa_forecast)
```
Basically as per Python. Since the best model is ARIMA(0,1,0) basically takes the last value to perpetuity

2.2 Auto arima with external factors

Incorporating various external factors


Define xreg list
```{r}
xreg_list <- c("Carbon.monoxide", "Humidity", "Nephelometer", "Nitric.Oxide", "Nitrogen.Dioxide", "Ozone", "PM10", "PM2.5", "Rainfall", 
                  "Sulphur.Dioxide", "Temperature", "Wind.Direction..10m.", "Wind.Direction.Sigma.Theta", "Wind.Speed..10m.")
```

Function that creates an xreg dataframe

```{r}
xreg_dataframe_builder <- function (df_set){

  xreg_datalist <- list()
  
  for (i in 1:length(xreg_list)){
    xreg_datalist[[xreg_list[i]]] <- df_set[, xreg_list[i]]
  }
  
  
  df <- do.call(cbind, xreg_datalist)
  
  return(df)
}
```

Build the xreg dataframes

```{r}
xreg_merriwa_train <- xreg_dataframe_builder(df_merriwa_train)
xreg_lidcombe_train <- xreg_dataframe_builder(df_lidcombe_train)

xreg_merriwa_test <- xreg_dataframe_builder(df_merriwa_test)
xreg_lidcombe_test <- xreg_dataframe_builder(df_lidcombe_test)

```

Train model with xreg

In Merriwa
```{r}

arima_merriwa_v2 <- auto.arima(df_merriwa_train[, 'CL_BLH'], xreg = xreg_merriwa_train)

checkresiduals(arima_merriwa_v2)
```

In Lidcombe

```{r}
arima_lidcombe <- auto.arima(df_lidcombe_train[, 'CL_BLH'], xreg = xreg_lidcombe_train)

checkresiduals(arima_lidcombe)
```



Forecasting

```{r}

forecasted_results_df <- function(arima_model, xreg_test, test_df){
  arima_forecast <- forecast(arima_model, h = 24, xreg = xreg_test)
  
  results_df <- data.frame(test_df[, 22:24])
  results_df['ARIMA model'] = arima_forecast$mean
  
  return(results_df)
}

```

Forecasted results

```{r}
results_df_merriwa_arima <- forecasted_results_df(arima_merriwa_v2, xreg_merriwa_test, df_merriwa_test)
results_df_lidcombe_arima <- forecasted_results_df(arima_lidcombe, xreg_lidcombe_test, df_lidcombe_test)
```


Stats

```{r}

stats_metrics(results_df_merriwa_arima$`ARIMA model`, results_df_merriwa_arima$CL_BLH, df_merriwa_train$CL_BLH, 'Merriwa stats')
```

```{r}

stats_metrics(results_df_lidcombe_arima$`ARIMA model`, results_df_lidcombe_arima$CL_BLH, df_lidcombe_train$CL_BLH, 'Lidcombe stats')
```

Plots

```{r}
predicted_plots <- function(forecasted_df, test_df, title){

  year <- as.integer(format(as.POSIXct(test_df$Date, format = "%Y-%m-%d"), format = "%Y"))
  month <- as.integer(format(as.POSIXct(test_df$Date, format = "%Y-%m-%d"), format = "%m"))
  date <- as.integer(format(as.POSIXct(test_df$Date, format = "%Y-%m-%d"), format = "%d"))
  
  forecasted_df$datetime <- ISOdatetime(year, month, date, df_merriwa_test[,'Hour'], 0, 0)
  
  pivot_df <- pivot_longer(forecasted_df, cols = 1:4, names_to = 'model', values_to = 'value')

  ggplot(pivot_df, aes(x = datetime, y = value, color = model))+
    geom_line(aes(linetype = model, color = model, size = model))+
    ylim(0,1800)+
    ggtitle(title)+
    scale_color_manual(values = c('red','blue','orange','green'))+
    scale_linetype_manual(values = c('solid','longdash','dashed','dashed'))+
    scale_size_manual(values = c(0.5,1,0.8,0.8))+
    theme(legend.title = element_blank())+
    labs(x = 'Date time', y = 'Boundary layer height (m)')
  
}

```


```{r}
predicted_plots(results_df_merriwa_arima, df_merriwa_test, 'PBL Height in Merriwa - ARIMA model')
```

```{r}
predicted_plots(results_df_lidcombe_arima, df_lidcombe_test, 'PBL Height in Lidcombe - ARIMA model')
```
Not bad.

A simple model - perhaps can be used as a alternative comaprison to the SARIMAX model + rolling 1 hour method developed in Python